---
title: "Inflation Report"
date:"r format(Sys.time(), '%d %B %Y')"
output: 
  html_document
    toc: yes
    toc_float:
      collapsed: no
      smooth_scrool: no
    toc_depth: 2
    df_print: paged
    theme: journal
  pdf_document:
    toc: yes
    toc_depth: '2'
--

The **Inflation Dashboard** shows the recent trends in inflation (HICP) and its components in the euro area.  
Data is sourced from Eurostat and the ECB.

Final estimate will be released on 17 April. April inflation estimate will be released on 30 April

For any queries, please contact **Aurelio Nocera (Monetary Policy Division)**.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align='center',echo = FALSE,warning = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

require("tidyr")
require("purrr")
require("dplyr")
require("xlsx")
require("zoo")
require("ggplot2")
require("plotly")
require("httr")
require("knitr")
require("kableExtra")
require("httr2")
library("lubridate")
library(glue)
library(curl)
library(flextable)
library(eurostat)

# tutorial https://ropengov.github.io/eurostat/articles/eurostat_tutorial.html#downloading-data

coicop_items= c("CP00","TOT_X_NRG_FOOD")

hicp = get_eurostat(id = "prc_hicp_manr", time_format = "date")
#hicp_ea_component = get_eurostat(id = "prc_hicp_ctrb", time_format = "date")

#Set proxies - VERY IMPORTANT IN THE BANK ELSE WILL NOT WORK
proxy_url <- "http://10.15.11.225:8080"
Sys.setenv(http_proxy = proxy_url, https_proxy = proxy_url, ftp_proxy = proxy_url)

euro_area = c("AT", "BE", "CY", "DE","EA","EE", "EL", "ES", "FI", "FR", "IE", "IT", "LU","LT","HR", 
       "LV", "MT", "NL", "PT", "SI", "SK","U2")

## Define your IAM credentials (Note this is Darwin log in details if different from SDW)
euaccount="IEANOCERA"
password="TempPass24%"

## Define format for data i.e. accept text or csv files from API
headers = c(`Accept` = 'text/csv')

# Combine the protocol wsEntryPoint and resource
source ="https://esb-ws-scde.eu.escb.eu/ws/SDW21/REST/service/data/ICP/"

##Headline SDW 

series_key<- "M..N.000000.4.ANR"
series_key= paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

hicp_sdw<- df[,c("TIME_PERIOD","OBS_VALUE","REF_AREA")]

hicp_sdw<- hicp_sdw%>%
  filter(REF_AREA %in% euro_area)

##Core SDW

series_key<- "M..N.XEF000.4.ANR"
series_key= paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

core_sdw<- df[,c("TIME_PERIOD","OBS_VALUE","REF_AREA")]

core_sdw<- core_sdw%>%
  filter(REF_AREA %in% euro_area)

###Headline

series_key <- "M.U2.Y.000000.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

hicpm_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

###Headline excl energy

series_key <- "M.U2.Y.XE0000.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

hicp5_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

###Core

series_key <- "M.U2.Y.XEF000.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

core_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

### Services

series_key <- "M.U2.Y.SERV00.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

services_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

###Goods

series_key <- "M.U2.Y.IGXE00.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

goods_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

### Foods

series_key <- "M.U2.Y.FOOD00.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

food_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

###NEIG

series_key <- "M.U2.Y.IGXE00.3.INX"
series_key = paste0(series_key) 
df <- httr::GET(glue(source, series_key), config(ssl_verifypeer = FALSE), authenticate(euaccount, password), 
                verbose(), add_headers(.headers = headers))           # Note that the target file is CSV format.
df= read.csv(textConnection(content(df, 'text')))

neig_sadj<- df[,c("TIME_PERIOD","OBS_VALUE")]

neig<- neig_sadj%>%
  mutate(momentum3m=4*((neig_sadj$OBS_VALUE-lag(neig_sadj$OBS_VALUE,3))/lag(neig_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(neig_sadj$OBS_VALUE-lag(neig_sadj$OBS_VALUE,12))/lag(neig_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

neig<-pivot_longer(neig,cols=2:3,names_to="Variable",values_to="Values")
neig<-neig%>%
  filter(Date>="2020-01")

write.csv(neig,"neig.csv")

### Transformation

core4<- core_sadj%>%
  mutate(momentum3m=4*((core_sadj$OBS_VALUE-lag(core_sadj$OBS_VALUE,3))/lag(core_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(core_sadj$OBS_VALUE-lag(core_sadj$OBS_VALUE,12))/lag(core_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

core4<-pivot_longer(core4,cols=2:3,names_to="Variable",values_to="Values")
core4<-core4%>%
  filter(Date>="2020-01")

write.csv(core4,"core_m_yoy.csv")

hicpm<- hicpm_sadj%>%
  mutate(momentum3m=4*((hicpm_sadj$OBS_VALUE-lag(hicpm_sadj$OBS_VALUE,3))/lag(hicpm_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(hicpm_sadj$OBS_VALUE-lag(hicpm_sadj$OBS_VALUE,12))/lag(hicpm_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

hicpm<-pivot_longer(hicpm,cols=2:3,names_to="Variable",values_to="Values")
hicpm<-hicpm%>%
  filter(Date>="2020-01")

write.csv(hicpm,"hicpm.csv")

hicp5<- hicp5_sadj%>%
  mutate(momentum3m=4*((hicp5_sadj$OBS_VALUE-lag(hicp5_sadj$OBS_VALUE,3))/lag(hicp5_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(hicp5_sadj$OBS_VALUE-lag(hicp5_sadj$OBS_VALUE,12))/lag(hicp5_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

hicp5<-pivot_longer(hicp5,cols=2:3,names_to="Variable",values_to="Values")
hicp5<-hicp5%>%
  filter(Date>="2020-01")

write.csv(hicp5,"hicp5.csv")

services<- services_sadj%>%
  mutate(momentum3m=4*((services_sadj$OBS_VALUE-lag(services_sadj$OBS_VALUE,3))/lag(services_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(services_sadj$OBS_VALUE-lag(services_sadj$OBS_VALUE,12))/lag(services_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

services<-pivot_longer(services,cols=2:3,names_to="Variable",values_to="Values")
services<-services%>%
  filter(Date>="2020-01")

write.csv(services,"services.csv")

goods<- goods_sadj%>%
  mutate(momentum3m=4*((goods_sadj$OBS_VALUE-lag(goods_sadj$OBS_VALUE,3))/lag(goods_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(goods_sadj$OBS_VALUE-lag(goods_sadj$OBS_VALUE,12))/lag(goods_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

goods<-pivot_longer(goods,cols=2:3,names_to="Variable",values_to="Values")
goods<-goods%>%
  filter(Date>="2020-01")

write.csv(goods,"goods.csv")

food<- food_sadj%>%
  mutate(momentum3m=4*((food_sadj$OBS_VALUE-lag(food_sadj$OBS_VALUE,3))/lag(food_sadj$OBS_VALUE,3)*100))%>%
  mutate(yoy_growth=100*(food_sadj$OBS_VALUE-lag(food_sadj$OBS_VALUE,12))/lag(food_sadj$OBS_VALUE,12))%>%
  mutate(Date=TIME_PERIOD)%>%
  select(Date,momentum3m,yoy_growth)

food<-pivot_longer(food,cols=2:3,names_to="Variable",values_to="Values")
food<-food%>%
  filter(Date>="2020-01")

write.csv(food,"food.csv")

```

# Flash estimate release

```{r}

hicp$TIME_PERIOD<- as.Date(paste(hicp$TIME_PERIOD,"01",sep="-"),format="%Y-%m-%d")

headline_euro_area = subset(hicp,geo %in% euro_area & TIME_PERIOD >="2020-01-01")%>%
 mutate(date=TIME_PERIOD)%>%
 select(coicop,geo,date,values)

headline_euro_area =subset(headline_euro_area,coicop =="CP00")

march<-max(headline_euro_area$date)
february<-max(headline_euro_area$date[headline_euro_area$date<march])

march_data<-headline_euro_area%>% filter(date==march)
february_data<- headline_euro_area%>% filter(date==february)

country_table<-data.frame(Country=unique(headline_euro_area$geo))

headline_result_table<- country_table%>%left_join(march_data%>%select(geo,March=values),by=c("Country"="geo"))%>%
 left_join(february_data %>% select(geo,February=values),by=c("Country"="geo"))

```


```{r}

core_euro_area = subset(hicp,geo %in% euro_area & TIME_PERIOD >="2020-01-01")%>%
 mutate(date=TIME_PERIOD)%>%
 select(coicop,geo,date,values)

core_euro_area =subset(core_euro_area,coicop =="TOT_X_NRG_FOOD")

march<-max(core_euro_area$date)
february<-max(core_euro_area$date[core_euro_area$date<march])

march_data<-core_euro_area%>% filter(date==march)
february_data<- core_euro_area%>% filter(date==february)

country_table<-data.frame(Country=unique(core_euro_area$geo))

core_result_table<- country_table%>%left_join(march_data%>%select(geo,March=values),by=c("Country"="geo"))%>%
 left_join(february_data %>% select(geo,February=values),by=c("Country"="geo"))

```

## Headline and Core 

```{r}

headline_core_table_2<- headline_result_table%>%left_join(core_result_table,by="Country")

colnames(headline_core_table_2)<- c("Country","Headline/Flash Mar","Headline/Feb","Core/Flash Mar","Core/Feb")

headline_core_table_2<- flextable(headline_core_table_2)

headline_core_table_2<-bold(headline_core_table_2,i=c(5,12),bold=TRUE,part="body")

headline_core_table_2

```

## Headline flash estimate

The March headline flash reading for euro area is 2.4%; March flash reading is above 2% in the majority of the euro area countries. 

```{r}

headline_euro_area = subset(hicp,geo %in% euro_area & TIME_PERIOD >=max(TIME_PERIOD))%>%
  mutate(date=TIME_PERIOD)%>%
  select(coicop,geo,date,values)

headline_euro_area =subset(headline_euro_area,coicop =="CP00")

headline_euro_area<- headline_euro_area%>%
  mutate(hicp_item= case_when(
    coicop == "CP00" ~ "Headline"))%>%
  select(hicp_item,date,geo,values)

headline_euro_area$geo<- factor(headline_euro_area$geo,levels=headline_euro_area$geo[order(headline_euro_area$values)])

chart3<-ggplot(headline_euro_area,aes(x=values,y=geo))+
  geom_bar(stat="identity",position="identity",fill="skyblue")+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

ggplotly(chart3)

```

(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

## Core flash estimate

The March core flash reading estimate is 2.9% for euro area; flash estimates for core are above 2% in all countries in the euro area.

```{r}

core_euro_area = subset(hicp,geo %in% euro_area & TIME_PERIOD >=max(TIME_PERIOD))%>%
  mutate(date=TIME_PERIOD)%>%
  select(coicop,geo,date,values)

core_euro_area =subset(core_euro_area,coicop =="TOT_X_NRG_FOOD")

core_euro_area<- core_euro_area%>%
  mutate(hicp_item= case_when(
    coicop == "TOT_X_NRG_FOOD" ~ "Core"))%>%
  select(hicp_item,date,geo,values)

core_euro_area$geo<- factor(core_euro_area$geo,levels=core_euro_area$geo[order(core_euro_area$values)])

chart4<-ggplot(core_euro_area,aes(x=values,y=geo))+
  geom_bar(stat="identity",position="identity",fill="skyblue")+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

ggplotly(chart4)

```
(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

## Euro Area

The March flash readings for inflation Headline and Core are 2.4% and 2.9%, respectively.

```{r}

### All items HICP, Food, 

hicp_core_ea<- hicp_sdw%>%
  left_join(core_sdw,by=c("TIME_PERIOD","REF_AREA"))%>%
  filter(REF_AREA =="U2")%>%
  filter(TIME_PERIOD>="2020-01")%>%
  select(TIME_PERIOD,OBS_VALUE.x,OBS_VALUE.y)

colnames(hicp_core_ea)<- c("Date","Headline","Core")

hicp_core_ea<-pivot_longer(hicp_core_ea,cols=2:3,names_to="Variable",values_to="Values")

hicp_core_ea$Date<- as.Date(paste(hicp_core_ea$Date,"01",sep="-"),format="%Y-%m-%d")

#hicp_ea = subset(hicp,geo=="EA" & time >="2020-01-01")%>%
#  mutate(date=time)%>%
#  select(coicop,geo,date,values)

#headline_core_ea =subset(hicp_ea,coicop %in% coicop_items)

#headline_core_ea<- headline_core_ea%>%
#  mutate(hicp_item= case_when(
#    coicop == "CP00" ~ "Headline",
#    coicop == "TOT_X_NRG_FOOD"~"Core"))%>%
#  select(hicp_item,date,geo,values)

chart1<-ggplot(hicp_core_ea,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("Headline"="#0083A0","Core"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="YoY growth rate, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart1)

```

(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

## Ireland

The March flash reading for the inflation Headline rate in Ireland is 1.7%. Core flash estimate for March is not yet available.

```{r}

### All items HICP, Food, 

hicp_core_ie<- hicp_sdw%>%
  left_join(core_sdw,by=c("TIME_PERIOD","REF_AREA"))%>%
  filter(REF_AREA =="IE")%>%
  filter(TIME_PERIOD>="2020-01")%>%
  select(TIME_PERIOD,OBS_VALUE.x,OBS_VALUE.y)

colnames(hicp_core_ie)<- c("Date","Headline","Core")

hicp_core_ie<-pivot_longer(hicp_core_ie,cols=2:3,names_to="Variable",values_to="Values")

hicp_core_ie$Date<- as.Date(paste(hicp_core_ie$Date,"01",sep="-"),format="%Y-%m-%d")

#hicp_ea = subset(hicp,geo=="EA" & time >="2020-01-01")%>%
#  mutate(date=time)%>%
#  select(coicop,geo,date,values)

#headline_core_ea =subset(hicp_ea,coicop %in% coicop_items)

#headline_core_ea<- headline_core_ea%>%
#  mutate(hicp_item= case_when(
#    coicop == "CP00" ~ "Headline",
#    coicop == "TOT_X_NRG_FOOD"~"Core"))%>%
#  select(hicp_item,date,geo,values)

chart2<-ggplot(hicp_core_ie,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("Headline"="#0083A0","Core"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="YoY growth rate, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart2)

```

(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

# HICP components 

## Headline

The March readings for the momentum and year on year growth rate of headline inflation are 3.91% and 2.40%, respectively.

```{r}

hicpm$Date<- as.Date(paste(hicpm$Date,"01",sep="-"),format="%Y-%m-%d")

chart20<-ggplot(hicpm,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart20)

```
Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## HICP excluding energy

The March readings for the momentum and year on year growth rate for HICP excluding energy are 3.30% and 2.85%, respectively.

```{r}

hicp5$Date<- as.Date(paste(hicp5$Date,"01",sep="-"),format="%Y-%m-%d")

chart6<-ggplot(hicp5,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart6)

```
Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## Core

The March readings for the momentum and year on year growth rate of core inflation are 3.72% and 2.88%, respectively.

```{r}

core4$Date<- as.Date(paste(core4$Date,"01",sep="-"),format="%Y-%m-%d")

chart5<-ggplot(core4,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart5)

```
Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## Services

The March readings for the momentum and year on year growth rate of services inflation are 5.29% and 3.89%, respectively.

```{r}

services$Date<- as.Date(paste(services$Date,"01",sep="-"),format="%Y-%m-%d")

chart7<-ggplot(services,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart7)
```

Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## Goods

The March readings for the momentum and year on year growth rate of goods inflation are 0.98% and 1.17%, respectively.

```{r}

goods$Date<- as.Date(paste(goods$Date,"01",sep="-"),format="%Y-%m-%d")

chart8<-ggplot(goods,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart8)

```

Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496))

## Food

The March readings for the momentum and year on year growth rate for food inflation are 1.78% and 2.71%, respectively.

```{r}

food$Date<- as.Date(paste(food$Date,"01",sep="-"),format="%Y-%m-%d")

chart9<-ggplot(food,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="3 months",date_labels="%b %Y")

ggplotly(chart9)

```

Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 



