---
title: "Inflation Report"
date: "`r format(Sys.time(),'%d %B %Y')`"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
    df_print: paged
    theme: journal
  pdf_document:
    toc: yes
    toc_depth: '2'
---

The **Inflation Report** shows the recent trends in inflation (HICP) and its components in the euro area.  Data is sourced from Eurostat and ECB.

August final estimates have been released on September 18. 
September flash estimates will be released on October 1st. 

For any queries, please do not hesitate to contact me. 
Author: **Aurelio Nocera**.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align='center',echo = FALSE,warning = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#Import ecb and eurostat libraries to connect to their APIs.
require(ecb)
require(eurostat)

# tutorial https://ropengov.github.io/eurostat/articles/eurostat_tutorial.html#downloading-data

coicop_items= c("CP00","TOT_X_NRG_FOOD")

hicp_eurostat = get_eurostat(id = "prc_hicp_manr", time_format = "date")
hicp_components_ea_eurostat = get_eurostat(id = "prc_hicp_ctrb", time_format = "date")

#Import packages to handle, manipulate and visualize data.

require("tidyr")
require("purrr")
require("dplyr")
require("xlsx")
require("zoo")
require("ggplot2")
require("plotly")
require("httr")
require("knitr")
require("kableExtra")
require("httr2")
require("lubridate")
require("glue")
require("curl")
require("lubridate")
require("zoo")
require("flextable")

# Define a list of euro area countries to import their inflation data.

euro_area = c("AT", "BE", "CY", "DE","EA","EE", "EL", "ES", "FI", "FR", "IE", "IT", "LU","LT","HR", "LV", "MT", "NL", "PT", "SI", "SK","U2")

## Headline, ECB - no seasonally adjusted

series_key<-"ICP.M..N.000000.4.ANR"
df<- ecb::get_data(series_key)

hicp_ea<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

##Core, ECB - no seasonally adjusted

series_key<- "ICP.M..N.XEF000.4.ANR"
df<- ecb::get_data(series_key)

core_ea<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

### Headline - ECB

series_key <- "ICP.M.U2.Y.000000.3.INX"
df<- ecb::get_data(series_key)

hicp_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

### Headline excl energy

series_key <- "ICP.M.U2.Y.XE0000.3.INX"
df<- ecb::get_data(series_key)

hicp_noenergy_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

### Core - ECB, sadj

series_key<- "ICP.M.U2.Y.XEF000.3.INX"
df<- ecb::get_data(series_key)

core_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

### Services - ECB, sadj

series_key <- "ICP.M.U2.Y.SERV00.3.INX"
df<- ecb::get_data(series_key)

services_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

### Goods - ECB, sadj

series_key <- "ICP.M.U2.Y.IGXE00.3.INX"
df<- ecb::get_data(series_key)

goods_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

### Foods - ECB, sadj

series_key <- "ICP.M.U2.Y.FOOD00.3.INX"
df<- ecb::get_data(series_key)

food_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

###NEIG - ECB, sadj

series_key <- "ICP.M.U2.Y.IGXE00.3.INX"
df<- ecb::get_data(series_key)

neig_sadj<- df%>%
  select(obstime,ref_area,obsvalue)%>%
  filter(ref_area %in% euro_area)

##Transformations for momentum charts 

neig<- neig_sadj%>%
  mutate(momentum3m=4*((neig_sadj$obsvalue-lag(neig_sadj$obsvalue,3))/lag(neig_sadj$obsvalue,3)*100))%>%
  mutate(yoy_growth=100*(neig_sadj$obsvalue-lag(neig_sadj$obsvalue,12))/lag(neig_sadj$obsvalue,12))%>%
  mutate(Date=obstime)%>%
  select(Date,momentum3m,yoy_growth)

neig<-pivot_longer(neig,cols=2:3,names_to="Variable",values_to="Values")
neig<-neig%>%
  filter(Date>="2020-01")

### Transformations: build momentum and yoy changes for each inflation series. 
### Time focus: 2020 onwards

hicp<- hicp_sadj%>%
  mutate(momentum3m=4*((hicp_sadj$obsvalue-lag(hicp_sadj$obsvalue,3))/lag(hicp_sadj$obsvalue,3)*100))%>%
  mutate(yoy_growth=100*(hicp_sadj$obsvalue-lag(hicp_sadj$obsvalue,12))/lag(hicp_sadj$obsvalue,12))%>%
  mutate(Date=obstime)%>%
  select(Date,momentum3m,yoy_growth)

hicp<-pivot_longer(hicp,cols=2:3,names_to="Variable",values_to="Values")
hicp<-hicp%>%
  filter(Date>="2020-01")

core<- core_sadj%>%
  mutate(momentum3m=4*((core_sadj$obsvalue-lag(core_sadj$obsvalue,3))/lag(core_sadj$obsvalue,3)*100))%>%
  mutate(yoy_growth=100*(core_sadj$obsvalue-lag(core_sadj$obsvalue,12))/lag(core_sadj$obsvalue,12))%>%
  mutate(Date=obstime)%>%
  select(Date,momentum3m,yoy_growth)

core<-pivot_longer(core,cols=2:3,names_to="Variable",values_to="Values")
core<-core%>%
  filter(Date>="2020-01")

services<- services_sadj%>%
  mutate(momentum3m=4*((services_sadj$obsvalue-lag(services_sadj$obsvalue,3))/lag(services_sadj$obsvalue,3)*100))%>%
  mutate(yoy_growth=100*(services_sadj$obsvalue-lag(services_sadj$obsvalue,12))/lag(services_sadj$obsvalue,12))%>%
  mutate(Date=obstime)%>%
  select(Date,momentum3m,yoy_growth)

services<-pivot_longer(services,cols=2:3,names_to="Variable",values_to="Values")
services<-services%>%
  filter(Date>="2020-01")

goods<- goods_sadj%>%
  mutate(momentum3m=4*((goods_sadj$obsvalue-lag(goods_sadj$obsvalue,3))/lag(goods_sadj$obsvalue,3)*100))%>%
  mutate(yoy_growth=100*(goods_sadj$obsvalue-lag(goods_sadj$obsvalue,12))/lag(goods_sadj$obsvalue,12))%>%
  mutate(Date=obstime)%>%
  select(Date,momentum3m,yoy_growth)

goods<-pivot_longer(goods,cols=2:3,names_to="Variable",values_to="Values")
goods<-goods%>%
  filter(Date>="2020-01")

food<- food_sadj%>%
  mutate(momentum3m=4*((food_sadj$obsvalue-lag(food_sadj$obsvalue,3))/lag(food_sadj$obsvalue,3)*100))%>%
  mutate(yoy_growth=100*(food_sadj$obsvalue-lag(food_sadj$obsvalue,12))/lag(food_sadj$obsvalue,12))%>%
  mutate(Date=obstime)%>%
  select(Date,momentum3m,yoy_growth)

food<-pivot_longer(food,cols=2:3,names_to="Variable",values_to="Values")
food<-food%>%
  filter(Date>="2020-01")

```

# Flash estimate release

```{r}

#hicp$obstime<- as.Date(paste(hicp$obstime,"01",sep="-"),format="%Y-%m-%d")

headline_euro_area = subset(hicp_eurostat,geo %in% euro_area & TIME_PERIOD >="2020-01-01")%>%
 mutate(date=TIME_PERIOD)%>%
 select(coicop,geo,date,values)

headline_euro_area =subset(headline_euro_area,coicop =="CP00")

august<-max(headline_euro_area$date)
july<-max(headline_euro_area$date[headline_euro_area$date<august])

august_data<-headline_euro_area%>% filter(date==august)
july_data<- headline_euro_area%>% filter(date==july)

country_table<-data.frame(Country=unique(headline_euro_area$geo))

headline_result_table<- country_table%>%left_join(august_data%>%select(geo,August=values),by=c("Country"="geo"))%>%
 left_join(july_data %>% select(geo,July=values),by=c("Country"="geo"))

```


```{r}

core_euro_area = subset(hicp_eurostat,geo %in% euro_area & TIME_PERIOD >="2020-01-01")%>%
 mutate(date=TIME_PERIOD)%>%
 select(coicop,geo,date,values)

core_euro_area =subset(core_euro_area,coicop =="TOT_X_NRG_FOOD")

august<-max(core_euro_area$date)
july<-max(core_euro_area$date[core_euro_area$date<august])

august_data<-core_euro_area%>% filter(date==august)
july_data<- core_euro_area%>% filter(date==july)

country_table<-data.frame(Country=unique(core_euro_area$geo))

core_result_table<- country_table%>%left_join(august_data%>%select(geo,August=values),by=c("Country"="geo"))%>%
 left_join(july_data %>% select(geo,July=values),by=c("Country"="geo"))

```

## Headline and Core 

```{r}

headline_core_table_2<- headline_result_table%>%left_join(core_result_table,by="Country")

colnames(headline_core_table_2)<- c("Country","Headline/Flash Aug","Headline/July","Core/Flash Aug","Core/July")

headline_core_table_2<- flextable(headline_core_table_2)

headline_core_table_2<-bold(headline_core_table_2,i=c(5),bold=TRUE,part="body")

headline_core_table_2

```

## Headline flash estimate

The June headline flash reading for euro area is 2.5%; June flash reading is still above 2% in the majority of the euro area countries. 

```{r}

headline_euro_area = subset(hicp_eurostat,geo %in% euro_area & TIME_PERIOD >=max(TIME_PERIOD))%>%
  mutate(date=TIME_PERIOD)%>%
  select(coicop,geo,date,values)

headline_euro_area =subset(headline_euro_area,coicop =="CP00")

headline_euro_area<- headline_euro_area%>%
  mutate(hicp_item= case_when(
    coicop == "CP00" ~ "Headline"))%>%
  select(hicp_item,date,geo,values)

headline_euro_area$geo<- factor(headline_euro_area$geo,levels=headline_euro_area$geo[order(headline_euro_area$values)])

chart1<-ggplot(headline_euro_area,aes(x=values,y=geo))+
  geom_bar(stat="identity",position="identity",fill="skyblue")+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

ggplotly(chart1)

```

(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

## Core flash estimate

The June core flash reading estimate is 2.9% for euro area; flash estimates for core are above 2% in the majority of all countries in the euro area.

```{r}

core_euro_area = subset(hicp_eurostat,geo %in% euro_area & TIME_PERIOD >=max(TIME_PERIOD))%>%
  mutate(date=TIME_PERIOD)%>%
  select(coicop,geo,date,values)

core_euro_area =subset(core_euro_area,coicop =="TOT_X_NRG_FOOD")

core_euro_area<- core_euro_area%>%
  mutate(hicp_item= case_when(
    coicop == "TOT_X_NRG_FOOD" ~ "Core"))%>%
  select(hicp_item,date,geo,values)

core_euro_area$geo<- factor(core_euro_area$geo,levels=core_euro_area$geo[order(core_euro_area$values)])

chart2<-ggplot(core_euro_area,aes(x=values,y=geo))+
  geom_bar(stat="identity",position="identity",fill="skyblue")+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

ggplotly(chart2)

```
(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

## Euro Area

The August readings for inflation Headline and Core are 2.2% and 2.8%, respectively.

```{r}

### All items HICP, Food, 

hicp_core_ea<- hicp_ea%>%
  left_join(core_ea,by=c("obstime","ref_area"))%>%
  filter(ref_area =="U2")%>%
  filter(obstime>="2020-01")%>%
  select(obstime,obsvalue.x,obsvalue.y)

colnames(hicp_core_ea)<- c("Date","Headline","Core")

hicp_core_ea<-pivot_longer(hicp_core_ea,cols=2:3,names_to="Variable",values_to="Values")

hicp_core_ea$Date<- as.Date(paste(hicp_core_ea$Date,"01",sep="-"),format="%Y-%m-%d")

chart3<-ggplot(hicp_core_ea,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("Headline"="#0083A0","Core"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="YoY growth rate, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="6 months",date_labels="%b %Y")

ggplotly(chart3)

```

(**Source: Eurostat** [link](https://ec.europa.eu/eurostat/databrowser/product/view/prc_hicp_manr))

# HICP components 

## Headline

The August readings for the momentum and year on year growth rate of headline inflation are 2.29% and 2.09%, respectively.

```{r}

hicp$Date<- as.Date(paste(hicp$Date,"01",sep="-"),format="%Y-%m-%d")

chart4<-ggplot(hicp,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="6 months",date_labels="%b %Y")

ggplotly(chart4)

```
Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## Core

The August readings for the momentum and year on year growth rate of core inflation are 2.93% and 2.74%, respectively.

```{r}

core$Date<- as.Date(paste(core$Date,"01",sep="-"),format="%Y-%m-%d")

chart5<-ggplot(core,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="6 months",date_labels="%b %Y")

ggplotly(chart5)

```
Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## Services

The August readings for the momentum and year on year growth rate of services inflation are 3.87% and 4.14%, respectively.

```{r}

services$Date<- as.Date(paste(services$Date,"01",sep="-"),format="%Y-%m-%d")

chart6<-ggplot(services,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="6 months",date_labels="%b %Y")

ggplotly(chart6)
```

Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 

## Goods

The August readings for the momentum and year on year growth rate of goods inflation are 1.24% and 0.34%, respectively.

```{r}

goods$Date<- as.Date(paste(goods$Date,"01",sep="-"),format="%Y-%m-%d")

chart7<-ggplot(goods,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="6 months",date_labels="%b %Y")

ggplotly(chart7)

```

Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496))

## Food

The August readings for the momentum and year on year growth rate for food inflation are 3.26% and 2.32%, respectively.

```{r}

food$Date<- as.Date(paste(food$Date,"01",sep="-"),format="%Y-%m-%d")

chart8<-ggplot(food,aes(x=Date,y=Values,color=Variable))+
  geom_line(size=1)+
  scale_color_manual(values=c("momentum3m"="#0083A0","yoy_growth"="#5EC5C2"))+
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())+
  labs(y="Percentage, %",color="")+
  theme(axis.text.x=element_text(size=12,angle=45,hjust=1),
        axis.text.y=element_text(size=12),
        legend.title=element_blank())+
  scale_x_date(date_breaks="6 months",date_labels="%b %Y")

ggplotly(chart8)

```

Momentum is defined as the annualised 3 months on 3 months rates, seasonally adjusted data.

(**Source: SDW** [link](https://sdw.escb.eu/browse.do?node=1496)) 
